### 2022-06-21 paper.Rmd
minor editing in example and discussion

Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups. This is due the diversity criterion being a measure of within-group heterogeneity; maximizing within-group heterogeneity only incidentally maximizes between-group similarity. However, because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity no longer holds unequal-sized groups. K-plus anticlustering strives for between-group similarity regardless of group size.

Unequal group sizes can be requested in anticlust by passing the different group sizes to the argument `K`. The following code was used to equalize means and variances between four groups using the k-plus approach:

@@

This function call relied on the default k-plus settings, ensuring that variances are equalized -- but no higher order moments or covariances.

@@

In this paper, I presented the k-plus criterion for anticlustering, which can be used to divided a data set into subgroups in such a way that the groups are similar to each other. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. The k-plus criterion also equalizes the variance between groups, and if needed any higher order moments such as skewness and kurtosis, or even the covariances. Thus, k-plus represents a family of anticlustering criteria, which can be combined by users according to their needs. Interestingly, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity.[^binarydata]  K-means anticlustering disregards any distribution characteristics other than the mean, while k-plus anticlustering ensures between-group similarity with regard to multiple criteria---and still minimizes differences in means quite well. The results showed that fulfilling several criteria at the same time may be aggregated in small data sets. Still, users can try out different objectives and see which specification fulfils their needs. A reasonable approach is to start greedy by requesting that multiple distribution moments should equalized. If the results of the greedy setup are subpar, users can reduce their requirements by opting for fewer criteria. Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.

[^binarydata]: A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.

@@

This paper focused on the presentation of new objective function but the did not really consider the multicriterion nature of the k-plus objective. Instead, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective. Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria. @brusco2019 showed am example of how direct algorithms can be used to optimize multiple criteria in anticlustering applications. Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.

@@

Focusing on the objective function only, and not on specialized algorithms, followed two reasonings. First, the applications showed that the local maximum search---maximizing an unweighted k-plus sum---yields satisfying results in practice that probably suffice most users. Second, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it. Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion regained attractivity: when opting for similarity in means and variances, k-plus anticlustering clearly outperformed the diversity criterion.

[^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion. The diversity criterion is the objective function used in the clustering method cluster editing, where it has to be minimized [@zahn1964; @shamir2004vf].

@@

This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the traditional k-means approach. K-plus anticlustering can be adopted whenever a data set has to be partitioned into equivalent parts. The method is accessible and easily usable via the R package `anticlust`.

>>>>

Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups. This is due the diversity criterion being a measure of within-group heterogeneity; maximizing within-group heterogeneity only incidentally maximizes between-group similarity. Because the diversity criterion is a sum rather than an average, the correspondence between within-group heterogeneity and between-group similarity does not hold for unequal-sized groups. K-plus anticlustering strives for between-group similarity regardless of group size.

Unequal group sizes can be requested in `anticlust` by passing the different group sizes to the argument `K`. The following code was used to equalize means and variances between four groups using the k-plus approach:

@@

This function call relied on the default k-plus settings ensuring similarity with regard to means and variances equalized---but with regard to higher order moments or covariances.

@@

In this paper, I presented the k-plus criterion for anticlustering, which can be used to divided a data set into subgroups in such a way that the groups are similar to each other. The k-plus criterion represents between-group similarity as discrepancies in distribution moments: mean, variances, skewness, kurtosis, any higher order moments, and covariances. Depending on the application, a different selection of these objectives can be included in the anticlustering process. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, this work indicates that k-plus anticlustering should be a default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be aggregated, especially in small data sets. In practice, users can try out different objectives and see which specification fulfils their needs. A reasonable approach is to start greedy by requesting that multiple distribution moments should equalized. If the results of such a greedy setup are subpar, users can reduce their requirements by opting for fewer criteria. Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering. Using the R package `anticlust`, application of k-plus anticlustering is easy and free.

[^binarydata]: A reasonable exception is given when the input data consists of binary variables because for binary data the mean and the variance are directly related.

@@

This paper focused on the presentation of new objective function but did not really consider the multicriterion nature of the k-plus criterion. Instead, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective. Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria. For example, @brusco2019 showed how direct algorithms can be used to optimize multiple criteria in anticlustering applications. Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion in applications that require high within-group heterogeneity.[^inanticlust] Future research should investigate how k-plus anticlustering may profit from similar---more sophisticated---algorithmic treatment.

@@

Focusing only on the objective function and not on optimization algorithms followed two reasonings. First, the applications showed that the local maximum search---maximizing the unweighted k-plus sum---yields satisfying results in practice that probably suffice most applications. Second, it was of interest to investigate the merits of the k-plus objective function per se, independently from the algorithm that is used to optimize it. In the simulation study, to obtain a fair assessment, it was important that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion. It was of particular interest to compare the k-plus objective with the common diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups. In contrast, maximizing the diversity leads to an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion has regained attractivity: when opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.

[^clusterediting]: In their study, they used the term anticluster editing to refer to the maximization of the diversity criterion. This is because the diversity criterion is minimized as part of the cluster editing method [@zahn1964; @shamir2004vf].

@@

This paper introduced k-plus anticlustering for maximizing between-group similarity, improving the classical k-means approach. K-plus anticlustering can be adopted whenever a group of elements has to be partitioned into equivalent parts. The method is accessible and easily usable via the R package `anticlust`.
